Classless Java

### 2015-11-15 paper/sections/Formalization.tex
Ready to read table text

What happens if we define a new interface \Q@M@ implementing one \Q@A@${}_i$ and one \Q@B@${}_i$? we have 9 cases, that can fit nicely a table:\\*

\begin{tabular}{|l|l|l|l|}
\hline
\textbf{M extends} & \textbf{A1}                  & \textbf{A2} & \textbf{A3} \\ \hline
\textbf{B1}        & error                        & error      &error       \\ \hline
\textbf{B2}        & accepted                        & error       & accepted       \\ \hline
\textbf{B3}        & \cellcolor[HTML]{C0C0C0}accepted &error       & error      \\ \hline
\end{tabular}
\\*
 and we have found that javac\footnote{We refer to \texttt{javac} as in JDK version is\marco{add}}
have a very strange behaviour in the following case

Table~\ref{table:javabug} shows the method overriding bug in Java. In the
example, there are 6 interfaces \texttt{A1,A2,A3,B1,B2,B3} with methods all
named as \texttt{m} which are either abstract or default methods. If we define a
new interface \texttt{M} that extends two of these interfaces, then the method
overriding result is shown in the table. For example, row2 col2 means \texttt{M
  extends A1,B1}. The result is \texttt{ERROR} because the abstract method
\texttt{T m();} in \texttt{A1} conflicts with the method \texttt{default T m()
  {...}} in \texttt{B1}. Readers may also figure out other extending cases by
following this way of interpretation. The interesting case in row4 col2 where
\texttt{M extends A1,B3}, the result should be \texttt{ERROR}, however in Java
the experimental result is \texttt{B3.m}. Because if we think consistently, this
case is esencially a method conflicting situation with abstract methods and
default methods, which is no different than the first case.

\lstinputlisting[linerange=5-10]{../UseMixinLombok/src/methodshadowing/test5/Test5.java} % APPLY:linerange=JAVABUG

In our approach, we overcome this bug by detecting method correctness when
typechecking all interfaces. Internally, algorithm in the auxiliary function
$\mBody(\m,\C)$ apply the correct method overriding mechanism and returns the
correct method signature.

>>>>

What happens if we define a new interface \Q@M@ implementing one \Q@A@${}_i$ and one \Q@B@${}_i$? we have 9 cases, that can fit nicely a table:\\*
\noindent
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{M extends} & \textbf{A1}                  & \textbf{A2} & \textbf{A3} \\ \hline
\textbf{B1}        & conservative error                        & conflict error      & conservative error       \\ \hline
\textbf{B2}        & both abstract, accepted                        & conservative error       & both abstract, accepted       \\ \hline
\textbf{B3}        & \cellcolor[HTML]{C0C0C0} accepted? &conflict  error       & conservative error      \\ \hline
\end{tabular}
\\*
We try to classify the results out of the table:
\begin{itemize}
\item \textbf{conflict error} happens when the method from \Q@A@${}_i$ and one \Q@B@${}_i$ are both implemented. This is also considered an error in most trait models.
\item \textbf{both abstract, accepted} happens when the method from \Q@A@${}_i$ and one \Q@B@${}_i$ are both abstract. This is also considered correct in all trait models.
\item \textbf{conservative error}
 happens when the method from \Q@A@${}_i$ and one \Q@B@${}_i$ is implemented in only one side.
The is different from what we would expect in a trait model, but it may make sense with the conservative idea that a method defined in an interface
should not silently satisfy a method in another one.
\end{itemize}

We are very confused on the result for \textbf{B3},\textbf{A1}, and we can not find any rational for this behaviour.
\footnote{We refer to the behaviour of \texttt{javac} as in JDK version is\marco{add}.}

%Table~\ref{table:javabug} shows the method overriding bug in Java. In the
%example, there are 6 interfaces \texttt{A1,A2,A3,B1,B2,B3} with methods all
%named as \texttt{m} which are either abstract or default methods. If we define a
%new interface \texttt{M} that extends two of these interfaces, then the method
%overriding result is shown in the table. For example, row2 col2 means \texttt{M
%  extends A1,B1}. The result is \texttt{ERROR} because the abstract method
%\texttt{T m();} in \texttt{A1} conflicts with the method \texttt{default T m()
%  {...}} in \texttt{B1}. Readers may also figure out other extending cases by
%following this way of interpretation.

% The interesting case in row4 col2 where
To be coherent with the idea of \textbf{conservative error}, the case should not be accepted.
We do not see how this should behave differently from \textbf{B1},\textbf{A1}, and \textbf{B3},\textbf{A3}.
We fear that the only retro-compatible fix for this strange behaviour is to accept all the cases of \textbf{conservative error} in a future version of Java.
%the experimental result is \texttt{B3.m}. Because if we think consistently, this

%\lstinputlisting[linerange=5-10]{../UseMixinLombok/src/methodshadowing/test5/Test5.java} % APPLY:linerange=JAVABUG

In our approach, we choose to not model this strange behaviour (a bug?).
Our auxiliary function $\mBody(\m,\C)$ enforce the \textbf{conservative error} strategy.
The rest of our formalization is parametric with the definition of $\mBody(\m,\C)$, thus if Java change its resolution strategy to a more permissive one, only minor adaptations in $\mBody(\m,\C)$ would be needed.
